Individual heterogeneity in the functional topography of the default-mode network in medial prefrontal cortex

### 2019-03-15 Manuscript.Rmd
second draft ready

---
title: 'Individual heterogeneity in the functional topography of the medial prefrontal cortex'
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  word_document: 
    reference_docx: Style_reference_draft.docx
  pdf_document: 
    keep_tex: true
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

  In this study, we aimed to subsection regions commonly attributed to both DMN and subjective value into subject-specific DMN and non-DMN partitions, quantifying the degree of topographical heterogeneity present in this network within and across individuals. We did this by capitalizing on the respective strengths of meta-analytic and subject-specific analyses of brain networks. First, we established a search space by selecting parcels from an established brain atlas [@Glasser2016] that considerably overlapped with previously-defined DMN and limbic networks along the medial wall [@Yeo2011]. A meta-analysis of DMN and SV literatures revealed areas in PCC and mPFC that were uniquely related to DMN and SV, respectively (these areas were used as markers to identify the affiliation of each individual's communities). We then produced rsFC networks of all the surface vertices within these regions of interest (ROI) for each of 100 individual resting state dataset from the Human Connectome Project (HCP)[@VanEssen2012], and estimated the extent of DMN coverage through spectral partitioning. We found that these areas of overlap can be partitioned into two communities that topographically vary across subjects, but that follow a comon organizational principle. This partition was in substantial agreement with results from a different community detection algorithm (modularity). A sliding window analysis showed that spectral partitioning allows for a more immediate identification of noisy viertices with unreliable affiliation. Refining the partitions by filtering out these nodes resulted in idiosyncratic partitions that were highly stable across scanning runs per individual, and this pattern was higher for mPFC. We show that partitions produced by spectral partitioning had higher test-retest reliability per subject than common seed-based functional connectivity, without the need of group-drived priors [@Kong2018]. Lastly, we provide evidence that these communities resembled an organizational scheme recently observed in a handful of subjects [@Braga2017]. Our work highlights the usefulness of estimating brain effects at the individual level in these regions, and points to limitations of aggregative methods in determining the degree of neural overlap between DMN and valuation.

@@

For all our analyses, we defined our search space based on the 17 network parcellation proposed by Yeo et al. [-@Yeo2011]. First, Ww selected vertices on the cortical surface that were contained by the DMN and limbic networks along the medial wall. Next, we overlayed these networks on a parcellated atlas of the human cortical surface (360 regions) [@Glasser2016], and retained parcels that shared at least 15% of vertices with them. This resulted in 45 regions of interest (ROI) across hemispheres (Table 1). These ROIs were naturally divided into two spatially discontiguous clusters that covered PCC and mPFC, thus facilitating the study of each general region separately.

@@

We used a novel approach to cortical surface parcel-based meta-analysis to assess whether individual parcels were differentially associated with valuation or DMN effects. For subjective value, we gathered peak brain coordinates of activity from 198 studies that were associated with positive effects for current or prospective rewards [@Bartra2013]. For DMN, we acquired coordinates from a previous meta-analysis of 80 studies that were related to task-deactivation [@Laird2009]. These represent the surviving areas post-statistical thresholding from each study. For each study, we created an indicator map in standard space (MNI152, 1 mm resolution) which contained values of 1 in a 10 mm radius sphere around each reported activation peak, and values of 0 elsewhere[@Wager2009]. The indicator map for each study was then projected to a standard cortical mesh (fs_average, 32,000 vertices, projfrac-max from 0 to 1 by 0.25) using FreeSurfer's mri_vol2surf [@Fischl2012] (http://surfer.nmr.mgh.harvard.edu/). Any sphere sections that overlapped with the cortex were mapped to our ROIs, and we tallied how many studies had positive indicator values in each parcel.

@@

Oue fMRI analyses used resting-state fMRI data from the Human Connectome Project [@VanEssen2012] Q6 release (n = `r nSubj`, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state runs (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) each comprised 1200 time points(14 min 33 s) for a totalof 4800 time points. Two runs were acquired on each day, with the phase encoding direction set to left-right for one run and right-left for the other. Only subjects with both left-right and right-left phase encoding were included (i.e. subjects with four rsfMRI sessions). In addition, only datasets with either low motion levels (under 1.5mm) or under 0.5 mm mean framewise displacement (FD) were used.

@@

Where $I$ is the identity matrix, $D$ is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlation values with every other vertex), and $W$ is the network weight matrix. This resulted in a matrix wherein each entry was the normalized value of the connection (from 0 to 1) between any two vertices relative to their combined connectivity strength, and with ones along the diagonal. This transformation ensures that every row adds up to zero. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (traditionally called the 'Fiedler vector') provides a set of positive and negative values to binarize the network with [@Fiedler1975]. This method enforces the partition of a network into similarly-sized communities, and guarantees that they share at least one edge [@Chung1999]. In this way, spectral partitioning avoids producing communities that are too small to be physiologically meaningful (for example, isolated vertices that are considerably affected by measurement noise). Every individual partition is therefore binarized by sign in order to evaluate the topography of each community (see next section for details). Given that this data-driven method does not label the two communities or establish correspondence across participants, we defined each individual's "DMN" community to be the one containing the DMN-specific PCC parcel identified in our meta-analysis. Given the high density of these networks (see Results), spectral partitioning was unlikely to face the issues associated with its use in sparse networks [@Fortunato2016].

In order to evaluate the convergent validity of the resulting partitions across community-detection methods, we also estimated these communities using the more traditional approach of modularity maximization, based on the algorithm from Clauset et al. [-@Clauset2004]. This method maximizes a quality function that compares the strength of the connection between any two vertices against a null model of their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combinations of vertices, and selects the clustering that maximizes the network's modularity. Unlike SP, modularity can fractionate a network into more than two functional groups. We assumed that if the partitions from the bounded and unbounded community detection methods showed high level of agreement, we could gain confidence in that we are indeed fractioning the ROIs into those belonging to DMN and not. We also diagnose our choice of two communities by reporting the rank (from lowest to largest) of the first eigenvalue produced by spectral partitioning to be non-zero. This eigenvalue serves as an estimator of the optimal number of communities in a network [@Chung1999], such that a rank of two in our partitions justifies limiting the number of communities to two.

@@

To estimate the degree of stability of these partitions per subject we performed a sliding window analysis (20 min windows, 1 min increments, mean number of windows per subject = 37), comparing each window's resulting partition against the partition derived from the whole subject's data. A 20-min window is typically adequate for identifying stable community features in brain networks [@Gordon2017]. We assessed whether the magnitude of the Fiedler vector value for a given vertex (for the full subject-level data set) was associated with the stability of that vertex’s sub-network assignment across time windows. To do this, we fit a mixed effects logistic regression model, in which the dependent variable was the proportion of times each vertex participated in the DMN community across windows, and the explanatory variables included a random effect of subject and a fixed effect of the Fiedler vector value for that vertex (derived from their full dataset). Based on this relationship, we identified a threshold of Fiedler vector values for each subject, such that vertices above these thresholds were persistently associated with either DMN or non-DMN over 95% of the time.

@@

We evaluated the performance of the spectral partitioning algorithm in comparison to a simpler partitioning approach using seed-based functional connectivity. First, we estimated each subject's DMN partition in mPFC based on its vertex-wise functional correlation (Pearson) with the average PCC activity for each day separately. We compared these seed-based maps with the Fiedler vectors produced by SP by spatially correlating them across days and subjects in three ways: 1) between seed-based maps; 2) between SP-based maps; and 3) between seed- and SP-based maps. These spatial correlations were aimed to determine the test/re-test reliability of each approach, as well as the overall level of agreement between them. The two methods are expected to produce somewhat similar arrangements, but the one displaying the higher within-subject agreement across days should be preferred (for a discussion on the stability of functional networks see [@Kong2018; @Mueller2013a]). We thus compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

@@

We performed a coordinate-based meta-analysis to identify unique standard regions across DMN and SV literatures. Volumetric coordinates from a set of DMN and SV studies were projected onto a cortical surface, and mapped to a multimodal cortical parcellation [@Glasser2016] to produce a list of brain areas reported per study. The 45 areas considered were limited to the medial portion of the default mode and limbic networks, as defined by the Yeo et al.[-@Yeo2011] 17 network parcellation. Unique areas were identified by first permuting the affiliation of each study (i.e. DMN or SV) to create a distribution of Chi-squared statistics based on the comparison of the number of reports per area between literatures (see Methods for details). This putative null distribution was used to identify regions more often reported in either literature.

Figure 1 shows the proportion of times every parcel was reported in each literature, as well as the resulting significant ROIs from the permutation. The 95th percentile of the permuted Chi-squared distribution was 9.11. Based on this threshold, area 7m was the only one to show unique bilateral affiliation to DMN (Left: observed $\chi^2$ = 10.07, *p* = 0.034; Right: observed $\chi^2$ = 18.89, *p* < 0.0001), while SV was uniquely represented bilaterally by area 25 (Left: observed $\chi^2$ = 12.91, *p* = 0.014; Right: observed $\chi^2$ = 12.83, *p* = 0.014). The purpose of these supra-threshold areas was to identify the most-likely affiliation of the two communities in our partitions by their coverage of these regions. However, area 25 sits in close proximity to the ventral striatum, which is reliably active in SV studies [@Bartra2013]. Considering that our projection from volumetric spheres to the cortex might have attributed striatal reports to this region, we decided not to use it as a marker of SV. We also chose not to use area v23 due to its unilateral significance (Right: observed $\chi^2$ = 11.51, *p* = 0.018). Instead, we only rely on area 7m to demarcate the DMN community.

@@

With these `r length(unique(str_remove(Yeo_labels, "_ROI")))` ROIs (`r (length(unique(str_remove(Yeo_labels, "_ROI"))) - 2)` overlapping, 1 DMN-specific, and 1 SV-specific), we estimated each individual's DMN coverage using their full time series (approximately 4800 total TRs from four scanning sessions over two days). Each subject's network (n = `r nSubj`) was constructed by performing pairwise correlations among the time series of all the surface vertices contained within the ROIs (`r nVertices` total vertices). After correcting for multiple comparisons, we performed spectral partitioning and modularity on these networks to identify cohesive functional communities denoting DMN and non-DMN areas. On average, networks had high densities (i.e. highly connected) regardless of the amount of data to produce them (mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`). This high degree of density is important, as these methods are known to face challenges when computed on sparse networks [@Fortunato2016].

Figure 2 shows a stereotypical subdivision of the ROIs for a single subject (100307) using spectral partitioning. This method dissects a network by the positive and negative values derived from the Fiedler vector (the eigenvector related to the second-to-lowest eigenvalue of the network's Laplacian matrix, see Methods), which we used to binarize the ROIs. Since this is a data-driven approach, there is no a priori labeling for these communities. We thus defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (7m), and ensured that the binary labeling of the vertices was consistent across subjects. In general, the resulting pattern shows large DMN coverage in PCC, with non-DMN vertices populating its anterior segment. On the other hand, mPFC is mostly covered by DMN vertices in its ventral-anterior and dorsal-anterior areas, with a persistent non-DMN pocket between them. This non-DMN section extends posteriorly onto pregenual cinglate cortex (area a24). Interstingly, this topographical combination of dorsal and ventral mPFC, with extensive PCC coverage, matches previous meta-analytic patterns more closely attributed to social cognition than subjective value [@Hiser2018].

Before determining the degree of generazibility of this topological pattern across individuals, we examined the validity of these partitions by comparing them to patterns derived with a modularity algorithm [@Clauset2004]. Modularity seeks to find the pattern of communities that maximizes a quality function that compares the connectivity strength of any two nodes with a null probability model. Since modularity is not bound to a desired number of partitions, it was capable of finding more than two communities in our datset. We quantified the cross-method agreement with the adjusted rand index (ARI). ARI measures the proportion of pairs of nodes in a network that were either clustered together or separately in both partitions, over all possible affiliation changes. This measure has the benefit of being agnostic to labeling schemes, as well as controlling for the expectation that the clusters were randomly produced [@Hubert1985]. This means that positive ARI values compare favorably against chance clustering.

@@

This comparison showed that both clustering methods were in high agreement across subjects (mean = `r round(mean(RI_all_within$Index), digits = 2)`, SD = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, max = `r max(nModularity)`). However, these communities encompassed a small number of vertices (median = `r median(modularityCounts_extra, na.rm = T)`, IQR = `r IQR(modularityCounts_extra, na.rm = T)`) compared to the principal two (median = `r median(modularityCounts_two, na.rm = T)`, IQR = `r IQR(modularityCounts_two, na.rm = T)`), suggesting that the overall network is composed of two main communities. In addition, the estimation of the optimal number of cuts based on the rank of the first non-zero eigenvalue (see "Partition evaluation" for details) showed that `r prcntDissection` percent of the individual networks are best partitioned into two communities.

Next, we examined how similar the partitions were across subjects by computing the ARI between every pair of subjects, and found that there was fairly low agreement overall (mean = `r round(mean(RI_all_between_lowtri), digits = 2)`, SD = `r round(sd(RI_all_between_lowtri), digits = 2)`). However, qualitative inspection of the brain community topologies showed good alignment for PCC, with a topologically similar mPFC pattern that shifted topographically across subjects. To quantify this heterogeneity in mPFC, we performed the same analysis for each region separately (Figure 3). Indeed, this analysis showed that the functional topology of PCC is better maintained across subjects (mean = `r round(mean(testPCC), digits = 2)`, SD = `r round(sd(testPCC), digits = 2)`), than mPFC (mean = `r round(mean(testPFC), digits = 2)`, SD = `r round(sd(testPFC), digits = 2)`)(paired permutation, *p* = `r Overall_PCCvsPFC_permTest$Pval`; Cohen's D = `r round(Overall_PCCvsPFC_ES, digits = 2)`).

@@

This analysis showed that the mean ARI along each subject's time series was high for both PCC (mean = `r round(mean(SC_PCCPFC_summary$Index[SC_PCCPFC_summary$Region=="PCC"]), digits = 2)`; SD = `r round(sd(SC_PCCPFC_summary$Index[SC_PCCPFC_summary$Region=="PCC"]), digits = 2)`) and mPFC (mean = `r round(mean(SC_PCCPFC_summary$Index[SC_PCCPFC_summary$Region=="mPFC"]), digits = 2)`; SD = `r round(sd(SC_PCCPFC_summary$Index[SC_PCCPFC_summary$Region=="mPFC"]), digits = 2)`), with significantly higher ARI for PCC (paired permutation,*p* = `r SW_PCCvPFC_permuteTest$Pval`; Cohen's D = `r round(SW_PCCvPFC_ES, digits = 2)`). These results suggest that the each individual's DMN distribution is cosiderably stable in time.

@@

To discard the possibility that the higher inter-subject variability in mPFC was driven by retaining vertices with lower signal quality, we compared the mean temporal signal to noise ratio (tSNR) for each region pre and post thresholding. We used subject time series that were not demeaned, but were otherwise equivalent to the data originally used. Paired permutation analyses (5000 iterations) showed that the mean tSNR in mPFC actually significantly increased post-thresholding across subjects (Pre: mean = `r snrR2_means$overall[1, 1]`, SD = `r snrR2_means$overall[3, 1]`; Post: mean = `r snrR2_means$overall[1, 2]`, SD = `r snrR2_means$overall[3, 2]`; paired permutation p-value = `r tsnrPermute_mPFC$Pval`, Cohen's D = `r tsnrCohen_mPFC[1]`), whereas the signal quality in PCC was preserved (Pre: mean = `r snrR2_means$overall[2, 1]`, SD = `r snrR2_means$overall[4, 1]`; Post: mean = `r snrR2_means$overall[2, 2]`, SD = `r snrR2_means$overall[4, 2]`; paired permutation p-value = `r tsnrPermute_PCC$Pval`, Cohen's D = `r tsnrCohen_PCC[1]`).

@@

The relative high variability seen in the aggregate map could be due to residual noise that was not captured by the sliding window analysis, or variations in the functional organization of mPFC across days. While the sliding window analysis suggested that these partitions are stable, many of the windows contained data from both days. A stricter evaluation of subject-specificity would be to compare every subject's subdivisions produced by data from each day separately. If the variability that we observed was truly due to having captured an individual's unique organization, day to day ARI values should be consistently higher within- than between-subjects. Figure 7 shows pairwise comparisons among ten example subjects for whole brain (left), as well as for PCC and mPFC separately (right).

@@

We can see again that the ratios for mPFC were higher than those of PCC (permutation,*p*= `r Sess_mPFCvPCC_permTest$Pval`; Cohen's D = `r round(Sess_mPFCvPCC_ES, digits = 2)`), but both were mostly above 1. This further confirms that the intrinsic functional organization of mPFC is uniquely arranged per individual, and provides evidence that such patterns can be identified with relatively small amounts of data in spite of a reduction in the estimates.

@@

### *Table 1. Regions of interest from Glasser et al. (2016). Prefixes indicate hemispheric location (e.g. R = right).*
| Literature           | Significant Brain Regions                 |
| :------------------- | ----------------------------------------: |
| ROIs                 | `r print(str_remove(Yeo_labels, "_ROI"))` |

>>>>

---
title: 'Individual heterogeneity in the functional topography of the default-mode network in medial prefrontal cortex'
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  word_document: 
    reference_docx: Style_reference_draft.docx
  pdf_document: 
    keep_tex: true
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

  In this study, we aimed to subsection regions commonly attributed to both DMN and subjective value into subject-specific DMN and non-DMN partitions, quantifying the degree of topographical heterogeneity present in this network within and across individuals. We did this by capitalizing on the respective strengths of meta-analytic and subject-specific analyses of brain networks. First, we established a search space by selecting parcels from an established brain atlas [@Glasser2016] that considerably overlapped with previously-defined DMN and limbic networks along the medial wall [@Yeo2011]. A meta-analysis of DMN and SV literatures revealed areas in PCC and mPFC that were uniquely related to DMN and SV, respectively (these areas were used as markers to identify the affiliation of each individual's communities). We then produced rsFC networks of all the surface vertices within these regions of interest (ROI) for each of 100 individual resting state dataset from the Human Connectome Project (HCP)[@VanEssen2012], and estimated the extent of DMN coverage through spectral partitioning. Focusing on individual vertices in this search space rather than the parcels (as is usual in brain network analyses) allowed us to finely delineate the topographical extent of each community. We found that these areas of overlap can be partitioned into two communities that topographically vary across subjects, but that follow a comon organizational principle. This partition was in substantial agreement with results from a different community detection algorithm (modularity). A sliding window analysis showed that spectral partitioning allows for a more immediate identification of noisy viertices with unreliable affiliation. Refining the partitions by filtering out these nodes resulted in idiosyncratic partitions that were highly stable across scanning runs per individual, and this pattern was higher for mPFC. We show that partitions produced by spectral partitioning had higher test-retest reliability per subject than common seed-based functional connectivity, without the need of group-drived priors [@Kong2018]. Lastly, we provide evidence that these communities resembled an organizational scheme recently observed in a handful of subjects [@Braga2017]. Our work highlights the usefulness of estimating brain effects at the individual level in these regions, and points to limitations of aggregative methods in determining the degree of neural overlap between DMN and valuation.

@@

For all our analyses, we defined our search space based on the 17 network parcellation proposed by Yeo et al. [-@Yeo2011]. First, We selected vertices on the cortical surface that were contained by the DMN and limbic networks along the medial wall. Next, we overlayed these networks on a parcellated atlas of the human cortical surface (360 regions) [@Glasser2016], and retained parcels that shared a considerable amount of vertices with them (visually inspected). This resulted in 45 regions of interest (ROI) across hemispheres (Table 1). These ROIs were naturally divided into two spatially discontiguous clusters that covered PCC and mPFC, thus facilitating the study of each general region separately.

@@

We used a novel approach to cortical surface parcel-based meta-analysis to assess whether individual parcels were differentially associated with valuation or DMN effects. For subjective value, we gathered peak brain coordinates of activity from 198 studies that were associated with positive effects for current or prospective rewards [@Bartra2013]. For DMN, we acquired coordinates from a previous meta-analysis of 80 studies that were related to task-deactivation [@Laird2009]. These coordinates represent the surviving areas post-statistical thresholding from each study. For each study, we created an indicator map in standard space (MNI152, 1 mm resolution) which contained values of 1 in a 10 mm radius sphere around each reported activation peak, and values of 0 elsewhere [@Wager2009]. The indicator map for each study was then projected to a standard cortical mesh (fs_average, 32,000 vertices, projfrac-max from 0 to 1 by 0.25) using FreeSurfer's mri_vol2surf [@Fischl2012] (http://surfer.nmr.mgh.harvard.edu/). Any sphere sections that overlapped with the cortex were mapped to our ROIs, and we tallied how many studies had positive indicator values in each parcel.

@@

Our fMRI analyses used resting-state fMRI data from the Human Connectome Project [@VanEssen2012] Q6 release (n = `r nSubj`, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state runs (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) each comprised 1200 time points(14 min 33 s) for a total of 4800 time points. Two runs were acquired on each day, with the phase encoding direction set to left-right for one run and right-left for the other. Only subjects with both left-right and right-left phase encoding were included (i.e. subjects with four rsfMRI sessions). In addition, only datasets with either low motion levels (under 1.5mm) or under 0.5 mm mean framewise displacement (FD) were used.

@@

Where $I$ is the identity matrix, $D$ is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlation values with every other vertex), and $W$ is the network weight matrix. This resulted in a matrix wherein each entry was the normalized value of the connection (from 0 to 1) between any two vertices relative to their combined connectivity strength, and with ones along the diagonal. This transformation ensures that every row adds up to zero. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (traditionally called the 'Fiedler vector') provides a set of positive and negative values to binarize the network with [@Fiedler1975]. This method enforces the partition of a network into similarly-sized communities, and guarantees that they share at least one edge [@Chung1999]. In this way, spectral partitioning avoids producing communities that are too small to be physiologically meaningful (for example, small sets of vertices that are spuriously correlated due to measurement noise). Every individual partition is therefore binarized by sign in order to evaluate the topography of each community (see next section for details). Given that this data-driven method does not label the two communities or establish correspondence across participants, we defined each individual's "DMN" community to be the one containing the DMN-specific PCC parcel identified in our meta-analysis. The high density of these networks (see Results) ensured that spectral partitioning did not face the issues associated with its use in sparse networks [@Fortunato2016].

In order to evaluate the convergent validity of the resulting partitions across community-detection methods, we also estimated these communities using the more traditional approach of modularity maximization, based on the algorithm from Clauset et al. [-@Clauset2004]. *(TBD) This method maximizes a quality function that compares the strength of the connection between any two vertices against a null model of their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combinations of vertices, and selects the clustering that maximizes the network's modularity.* Unlike SP, modularity can fractionate a network into more than two functional groups. We assumed that if the partitions from the bounded and unbounded community detection methods showed high levels of agreement, we could gain confidence in that we are indeed fractioning the ROIs into those belonging to DMN and not. We also diagnosed our choice of two communities by reporting the rank (from lowest to largest) of the first eigenvalue produced by spectral partitioning to be non-zero. This eigenvalue serves as an estimator of the optimal number of communities in a network [@Chung1999], such that a rank of two in our partitions justifies limiting the number of communities to two.

@@

To estimate the degree of stability of these partitions per subject we performed a sliding window analysis (20 min windows, 1 min increments, mean number of windows per subject = 37), comparing each window's resulting partition against the partition derived from the whole subject's data. A 20-min window is typically adequate for identifying stable community features in brain networks [@Gordon2017]. We assessed whether the magnitude of the Fiedler vector value for a given vertex (for the full subject-level data set) was associated with the stability of that vertex’s sub-network assignment across time windows. To do this, we fit a mixed effects logistic regression model, in which the dependent variable was the proportion of times each vertex participated in the DMN community across windows, and the explanatory variables included a random effect of subject and a fixed effect of the Fiedler vector value for that vertex (derived from their full time series). Based on this relationship, we identified a threshold of Fiedler vector values for each subject, such that vertices above these thresholds were persistently associated with either DMN or non-DMN over 95% of the time.

@@

We evaluated the performance of the spectral partitioning algorithm in comparison to a simpler partitioning approach using seed-based functional connectivity. First, we estimated each subject's DMN partition in mPFC based on its vertex-wise functional correlation (Pearson) with the average PCC activity for each day separately. We compared these seed-based maps with the Fiedler vectors produced by SP by spatially correlating them across days and subjects in three ways: 1) between seed-based maps; 2) between SP-based maps; and 3) between seed- and SP-based maps. These spatial correlations were aimed to determine the test/re-test reliability of each approach, as well as the overall level of agreement between them. The two methods were expected to produce somewhat similar arrangements, but the one displaying the higher within-subject agreement across days should be preferred (for a discussion on the stability of functional networks see [@Kong2018; @Mueller2013a]). We thus compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

@@

We performed a coordinate-based meta-analysis to identify unique standard regions across DMN and SV literatures. Volumetric coordinates from a set of DMN and SV studies were projected onto a cortical surface, and mapped to a multimodal cortical parcellation [@Glasser2016] to produce a list of brain areas reported per study. The 45 areas considered were limited to the medial portion of the default mode and limbic networks, as defined by the Yeo et al.'s [-@Yeo2011] 17 network parcellation. Unique areas were identified by first permuting the affiliation of each study (i.e. DMN or SV) to create a distribution of chi-squared statistics based on the comparison of the number of reports per area between literatures (see Methods for details). This putative null distribution was used to identify regions more often reported in either literature.

Figure 1 shows the proportion of times every parcel was reported in each literature, as well as the resulting significant ROIs from the permutation. The 95th percentile of the permuted chi-squared distribution was 9.11. Based on this threshold, area 7m was the only one to show unique bilateral affiliation to DMN (Left: observed $\chi^2$ = 10.07, *p* = 0.034; Right: observed $\chi^2$ = 18.89, *p* < 0.0001), while SV was uniquely represented bilaterally by area 25 (Left: observed $\chi^2$ = 12.91, *p* = 0.014; Right: observed $\chi^2$ = 12.83, *p* = 0.014). The purpose of these supra-threshold areas was to identify the most-likely affiliation of the two communities in our partitions by their coverage of these regions. However, area 25 sits in close proximity to the ventral striatum, which is reliably active in SV studies [@Bartra2013]. Considering that our projection from volumetric spheres to the cortex might have attributed striatal reports to this region, we decided not to use it as a marker of SV. We also chose not to use area v23 due to its unilateral significance (Right: observed $\chi^2$ = 11.51, *p* = 0.018). Instead, we only rely on area 7m to demarcate the DMN community.

@@

With these `r length(unique(str_remove(Yeo_labels, "_ROI")))` ROIs (`r (length(unique(str_remove(Yeo_labels, "_ROI"))) - 2)` overlapping, 1 DMN-specific, and 1 SV-specific), we estimated each individual's DMN coverage using their full time series (approximately 4800 total TRs from four scanning sessions over two days). Each subject's network (n = `r nSubj`) was constructed by performing pairwise correlations among the time series of all the surface vertices contained within the ROIs (`r nVertices` total vertices). After correcting for multiple comparisons, we performed spectral partitioning on these networks to identify cohesive functional communities denoting DMN and non-DMN areas. On average, networks had high densities (i.e. highly connected) regardless of whether they were produced with a single run all the full time series (mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`). This high degree of density is important, as this method is known to face challenges when computed on sparse networks [@Fortunato2016].

Figure 2 shows a stereotypical subdivision of the ROIs for a single subject (100307) using spectral partitioning. This method dissects a network by the positive and negative values derived from the Fiedler vector (the eigenvector related to the second-to-lowest eigenvalue of the network's Laplacian matrix, see Methods), which we used to binarize the ROIs. Since this is a data-driven approach, there is no a priori labeling for these communities. We thus defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (7m), and ensured that the binary labeling of the vertices was consistent across subjects. In general, the resulting pattern shows large DMN coverage in PCC, with non-DMN vertices populating its anterior segment. On the other hand, mPFC is mostly covered by DMN vertices in its ventral-anterior and dorsal-anterior areas, with a persistent non-DMN pocket between them. This non-DMN section extends posteriorly onto pregenual cinglate cortex (area a24).

Before determining the degree of generazibility of this topological pattern across individuals, we examined the validity of these partitions by comparing them to patterns derived with a modularity algorithm [@Clauset2004]. Modularity seeks to find the pattern of communities that maximizes a quality function that compares the connectivity strength of any two nodes with a null probability model. Since modularity is not bound to a desired number of communities, it was capable of finding more than two in our datset. We quantified the cross-method agreement with the adjusted rand index (ARI). ARI measures the proportion of pairs of nodes in a network that were either clustered together or separately in both partitions, over all possible affiliation changes. This measure has the benefit of being agnostic to labeling schemes, as well as controlling for the expectation that the clusters were randomly produced [@Hubert1985]. This means that positive ARI values compare favorably against chance clustering.

@@

This comparison showed that both clustering methods were in high agreement across subjects (mean = `r round(mean(RI_all_within$Index), digits = 2)`, SD = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, max = `r max(nModularity)`). However, these communities encompassed a small number of vertices (mean = `r mean(modularityCounts_extra, na.rm = T)`, IQR = `r IQR(modularityCounts_extra, na.rm = T)`) compared to the principal two (mean = `r mean(modularityCounts_two, na.rm = T)`, IQR = `r IQR(modularityCounts_two, na.rm = T)`), suggesting that the overall network is composed of two main communities. In addition, the estimation of the optimal number of cuts based on the rank of the first non-zero eigenvalue (see "Partition evaluation" for details) showed that `r prcntDissection` percent of the individual networks are best partitioned into two communities.

Next, we examined how similar the partitions were across subjects by computing the ARI between every pair of subjects, and found that there was moderate agreement overall (mean = `r round(mean(RI_all_between_lowtri), digits = 2)`, SD = `r round(sd(RI_all_between_lowtri), digits = 2)`). However, qualitative inspection of the brain community topologies showed good alignment for PCC, with a topologically similar mPFC pattern that shifted topographically across subjects. To quantify this heterogeneity in mPFC, we performed the same analysis for each region separately (Figure 3). Indeed, this analysis showed that the functional topology of PCC is better maintained across subjects (mean = `r round(mean(testPCC), digits = 2)`, SD = `r round(sd(testPCC), digits = 2)`) than mPFC (mean = `r round(mean(testPFC), digits = 2)`, SD = `r round(sd(testPFC), digits = 2)`)(paired permutation, *p* = `r Overall_PCCvsPFC_permTest$Pval`; Cohen's D = `r round(Overall_PCCvsPFC_ES, digits = 2)`).

@@

This analysis showed that the mean ARI along each subject's time series was high for both PCC (mean = `r round(mean(SC_PCCPFC_summary$Index[SC_PCCPFC_summary$Region=="PCC"]), digits = 2)`; SD = `r round(sd(SC_PCCPFC_summary$Index[SC_PCCPFC_summary$Region=="PCC"]), digits = 2)`) and mPFC (mean = `r round(mean(SC_PCCPFC_summary$Index[SC_PCCPFC_summary$Region=="mPFC"]), digits = 2)`; SD = `r round(sd(SC_PCCPFC_summary$Index[SC_PCCPFC_summary$Region=="mPFC"]), digits = 2)`), with significantly higher ARI for PCC (paired permutation, *p* = `r SW_PCCvPFC_permuteTest$Pval`; Cohen's D = `r round(SW_PCCvPFC_ES, digits = 2)`). These results suggest that the each individual's DMN distribution is cosiderably stable in time.

@@

To discard the possibility that the higher inter-subject variability in mPFC was driven by retaining vertices with lower signal quality, we compared the mean temporal signal to noise ratio (tSNR) for each region pre and post thresholding. We used subject time series that were not demeaned, but were otherwise equivalent to the data originally used. Paired permutation analyses (5000 iterations) showed that the mean tSNR in mPFC actually significantly increased post-thresholding across subjects (Pre: mean = `r snrR2_means$overall[1, 1]`, SD = `r snrR2_means$overall[3, 1]`; Post: mean = `r snrR2_means$overall[1, 2]`, SD = `r snrR2_means$overall[3, 2]`; paired permutation p-value = `r tsnrPermute_mPFC$Pval`, Cohen's D = `r tsnrCohen_mPFC[1]`), whereas the mean signal quality in PCC was unaffected (Pre: mean = `r snrR2_means$overall[2, 1]`, SD = `r snrR2_means$overall[4, 1]`; Post: mean = `r snrR2_means$overall[2, 2]`, SD = `r snrR2_means$overall[4, 2]`; paired permutation p-value = `r tsnrPermute_PCC$Pval`, Cohen's D = `r tsnrCohen_PCC[1]`).

@@

The relative high variability seen in the aggregate map could be due to variations in the functional organization of mPFC across days. While the sliding window analysis suggested that these partitions are stable, many of the windows contained data from both days. A stricter evaluation of subject-specificity would be to compare every subject's subdivisions produced by data from each day separately. If the variability that we observed was truly due to having captured an individual's unique organization, day to day ARI values should be consistently higher within- than between-subjects. Figure 7 shows pairwise comparisons among ten example subjects for whole brain (left), as well as for PCC and mPFC separately (right).

@@

We can see again that the ratios for mPFC were higher than those of PCC (permutation,*p*= `r Sess_mPFCvPCC_permTest$Pval`; Cohen's D = `r round(Sess_mPFCvPCC_ES, digits = 2)`), but both were mostly above 1. This further confirms that the intrinsic functional organization of mPFC is uniquely arranged per individual, and provides evidence that such patterns can be identified with small amounts of data in spite of a reduction in the estimates.

@@

### *Table 1. Regions of interest from Glasser et al. (2016). Prefixes indicate hemispheric location (e.g. R = right).*
| Literature           | Significant Brain Regions                 |
| :------------------- | ----------------------------------------: |
| ROIs                 | `r str_remove(Yeo_labels, "_ROI")` |
